\vspace{-0.25em}\thetitle

### 2009-07-29 paper.tex
...

We use randomly sampled traffic from six different network traces.
All six traces are freely available from the \caps{CRAWDAD} wireless trace repository~\cite{Yeo06}.
Details of the traces are shown in \Table{traces}.
These traces represent a broad cross-section of traffic patterns over time and a reasonable variety of network types.
Unfortunately, we could not find and freely available commercial hotspot or corporate network traces that were usable for our analysis.
We randomly sampled 5000 flows from each trace for training and another 5000 flows for testing.

@@

Once the data have been expressed in matrix form, the next step is attempt to detect the linear structures seen in \Figure{svd}.
% Recall that these correspond to the point classes in our hypothesized model.
We use Ma~\emph{et~al.}'s~\cite{Ma07} algorithm for segmenting multivariate data into a union of subspaces using lossy data coding and compression.
The essential idea of this algorithm is to compute the number of bits needed to encode the data under various assumed groupings.
The grouping which yields the best compression is the desired segmentation.
This algorithm has many desirable qualities.
Two are essential:
1) unlike most clustering algorithms, it finds subspaces, not regions;
2) it is highly robust to outliers and noisy data.
The algorithm has one parameter:
the maximum allowable distortion of the data, $\varepsilon^2$.
We follow Ma's lead when applying the algorithm to \caps{SVD} data and use $\varepsilon^2 = 1$.
The result of this stage is a segmentation of the data into classes of rows, precisely as our hypothesized model requires.

@@

\newfootnote{\svdnnnote}{The first singular vector of a nonnegative matrix is guaranteed to be nonnegative, but the next will generally not be.}

@@

Our technique is fundamentally similar to fitting a regression line to noisy two-dimensional training data and then using the $x$-coordinates of test data points to linearly predict the $y$-coordinates.
The main differences are the vastly higher number of dimensions and more complex structure we assume for our data.
From the training data, we recover our estimated version of the $P$ matrix.
We split this into its observable and predictable parts: $P_\text{o}$ and $P_\text{p}$.
Given an observation matrix, $X^*_\text{o}$, to perform prediction, we find the nonnegative matrix of weights, $W_*$, minimizing the squared Frobenius error:
\begin{align}
  \norm{X^*_\text{o} - W^* P_\text{o}}^2_\text{frob}.
\end{align}
This is equivalent to minimizing the square error for each observation individually, but optimizing \emph{en masse} allows fast non-negative least squares algorithms to be used~\cite{Benthem04,Kim08:block-pivot}.
The predicted behavior for these flows is simply weight matrix times the predictable feature matrix:
\begin{align}
  X^*_\text{p} = W^* P_\text{p}.
\end{align}
The rows of $X^*_\text{p}$ are our prediction about the behavior of the flows observed in $X^*_\text{o}$.
Each row contains a packet size distribution, an inter-packet interval distribution and a distribution of packet counts.

@@

\subsection{Evaluating Predictions}\sec{evaluation}

@@

\begin{figure*}[t]
\begin{center}
\subfloat[Packet sizes: \footnotesize{DARTMOUTH}]{%
\includegraphics[width=3.545in]{size/dart}}
\subfloat[Packet sizes: \footnotesize{SIGCOMM 2004}]{%
\includegraphics[width=3.545in]{size/sc04}}
\subfloat[Inter-packet intervals: \footnotesize{DARTMOUTH}]{%
\includegraphics[width=3.545in]{ival/dart}}
\subfloat[Inter-packet intervals: \footnotesize{SIGCOMM 2004}]{%
\includegraphics[width=3.545in]{ival/sc04}}
\caption{%
Empirical CDFs of Kolmogorov-Smirnov p-values for various predictors of packet size distribution and inter-packet interval distribution for two selected traces. Lower curves are better and the ideal curve provides a lower bound on quality.
Vertical segments indicate the farthest distance between the ideal and each CDF.
This distance serves as a pessimistic measure of each predictor's error rate.
}
\fig{cdf}
\end{center}
\vspace{-2em}
\end{figure*}

>>>>

We use randomly sampled traffic from six different network traces.
All six traces are freely available from the \caps{CRAWDAD} wireless trace repository~\cite{Yeo06}.
Details of the traces are shown in \Table{traces}.
These traces represent a broad cross-section of traffic patterns over time and a reasonable variety of network types.
Unfortunately, we could not find any freely available residential or corporate network traces that were usable for our analysis.
We randomly sampled 5000 flows from each trace for training and another 5000 flows for testing.

@@

Once the data have been expressed in matrix form, the next step is attempt to detect the linear structures seen in \Figure{svd}.
% Recall that these correspond to the point classes in our hypothesized model.
We use Ma~\emph{et~al.}'s~\cite{Ma07} algorithm for segmenting multivariate data into a union of subspaces using lossy data coding and compression.
The essential idea of this algorithm is to compute the number of bits needed to encode the data under various assumed groupings.
The grouping which yields the best compression is the desired segmentation.
This algorithm has many desirable qualities.
Two are essential:
1) unlike most clustering algorithms, it finds subspaces, not regions;
2) it is highly robust to outliers and noisy data.
The algorithm has one parameter:
the maximum allowable distortion of the data
We follow Ma's lead when applying the algorithm to \caps{SVD} data and use $\varepsilon^2 = 1$.
The result of this stage is a segmentation of the data into classes of rows, precisely as our hypothesized model requires.

@@

\newfootnote{\svdnnnote}{The first singular vector of a nonnegative matrix is guaranteed to be nonnegative, but the other singular vectors will have mixed sign.}

@@

Our technique is fundamentally similar to fitting a regression line to noisy two-dimensional training data and then using the $x$-coordinates of test data points to linearly predict the $y$-coordinates.
The main differences are the vastly higher number of dimensions and more complex structure we assume for our data.
From the training data, we recover our estimated version of the $P$ matrix.
We split this into its observable and predictable parts: $P_\text{o}$ and $P_\text{p}$.
Given an observation matrix, $X^*_\text{o}$, to perform prediction, we find the nonnegative matrix of weights, $W_*$, minimizing the squared Frobenius error:
\begin{align}
  \norm{X^*_\text{o} - W^* P_\text{o}}^2_\text{frob}.
\end{align}
This is equivalent to minimizing the square error for each observation individually, but optimizing \emph{en masse} allows fast non-negative least squares algorithms to be used~\cite{Benthem04,Kim08:block-pivot}.
The predicted behavior for these flows is simply the weight matrix times the predictable feature matrix:
\begin{align}
  X^*_\text{p} = W^* P_\text{p}.
\end{align}
The rows of $X^*_\text{p}$ are our predictions about the behavior of flows observed in $X^*_\text{o}$.
Each row contains a packet size distribution, an inter-packet interval distribution and a distribution of packet counts.

@@

@@

\begin{figure*}[t]
\vspace{-1em}
\begin{center}
\subfloat[Packet sizes: \footnotesize{DARTMOUTH}]{%
\includegraphics[width=3.545in]{size/dart}}
\subfloat[Packet sizes: \footnotesize{SIGCOMM 2004}]{%
\includegraphics[width=3.545in]{size/sc04}}
\subfloat[Inter-packet intervals: \footnotesize{DARTMOUTH}]{%
\includegraphics[width=3.545in]{ival/dart}}
\subfloat[Inter-packet intervals: \footnotesize{SIGCOMM 2004}]{%
\includegraphics[width=3.545in]{ival/sc04}}
\caption{%
Empirical CDFs of Kolmogorov-Smirnov p-values for various predictors of packet size distribution and inter-packet interval distribution for two selected traces. Lower curves are better and the ideal curve provides a lower bound on quality.
Vertical segments indicate the farthest distance between the ideal and each CDF.
This distance serves as a pessimistic measure of each predictor's error rate.
}
\fig{cdf}
\end{center}
\vspace{-1em}
\end{figure*}
